Hybrid Threads for the Parrot Virtual Machine

### 2012-05-12 Paper.tex

\section{Erlang}

@@

Contrary to other widely used virtual machines like the JVM or the CLR which are stack based, Parrot mirrors contemporary hardware CPUs more closely by being register based. A stack based virtual machine usually fetches the operands of an operation from the top of a stack and puts the result back on top. Thus the operands chosen implicitly by ordering of the operations allowing the opcodes to be quite small. In a register based VM on the other hand each operation has to specify the operands explicitly. Compilers for stack machines are simpler because they do not have to care about register allocations and code is independent of prior or subsequent code\cite{VMShowdown}. The rationale behind giving up the simplicity of a stack based implementation is the hope of simplifying just in time compilation and improved performance of nested function and method calls.

@@

PMCs (PolyMorphic Container) are containers for all higher level types such as objects, arrays, hash tables or code. Thus they are very similar to Python's PythonObject types. PMCs are defined by C structs and are fully garbage collected. Their definition looks like:
\begin{CCode}
struct PMC {
    Parrot_UInt    flags;
    VTABLE         *vtable;             /* Pointer to vtable. */
    DPOINTER       *data;               /* Pointer to attribute structure. */
    PMC            *_metadata;          /* Pointer to metadata PMC. */
};
\end{CCode}
In short it contains a pointer to a type specific data structure and some meta data. The vtable is where the type's behavior is defined. It contains a long list of function pointers forming Parrot's unified type interface. This interface is a union of numeric, string, array, hash and object like behaviors. For example the  \textit{get\_integer} function returns an integer value for the data type. For a simple \textit{int} it is the contained value. For an array it may be the number of elements. The \textit{find\_method} function makes user defined methods of objects possible.

@@

Control flow is modeled in Parrot using continuation passing. A continuation is a data structure containing the state of a program at a given point in its execution. It contains all information needed to continue a program in a certain state, e.g. a call stack, the instruction pointer and contents of local variables. ``Calling'' a continuation means restoring the state encapsulated in the continuation.

@@

A runloop is in principle the most inner loop executing bytecode in Parrot. It is also a data structure containing some information needed to support exception handling. When some operation in Parrot bytecode is calling a C function and this C function in turn is again executing Parrot bytecode, a nested runloop is started. Examples for such situations include calling a library function with a callback as parameter and exceptions thrown inside Parrot's C code calling a previously defined exception handler in user code.

@@

\textit{Parrot\_cx\_schedule\_sleep} is the actual implementation of the \textit{sleep} op. Like in\textit{Parrot\_cx\_preempt\_task}, \textit{Parrot\_cx\_stop\_task} is used to get an updated task at the current execution position, but instead of pushing this on the run queue, it is used as a callback for a newly set alarm. Again the NULL opcode is used to stop processing of the current task.

>>>>

@@

Contrary to other widely used virtual machines like the JVM or the CLR which are stack based, Parrot mirrors contemporary hardware CPUs more closely by being register based. A stack based virtual machine usually fetches the operands of an operation from the top of a stack and puts the result back on top. Thus the operands are chosen implicitly by ordering of the operations allowing the opcodes to be quite small. In a register based VM on the other hand each operation has to specify the operands explicitly. Compilers for stack machines are simpler because they do not have to care about register allocations and code is independent of prior or subsequent code\cite{VMShowdown}. The rationale behind giving up the simplicity of a stack based implementation is the hope of simplifying just in time compilation and improved performance of nested function and method calls.

@@

PMCs (PolyMorphic Container) are containers for all higher level types such as objects, arrays, hash tables or code. Thus they are very similar to Python's PythonObject types. PMCs are defined by C structs and are fully garbage collected. Their definition looks like:
\begin{CCode}
struct PMC {
    Parrot_UInt    flags;
    VTABLE         *vtable;             /* Pointer to vtable. */
    DPOINTER       *data;               /* Pointer to attribute structure. */
    PMC            *_metadata;          /* Pointer to metadata PMC. */
};
\end{CCode}
In short they contain a pointer to a type specific data structure and some meta data. The vtable is where the type's behavior is defined. It contains a long list of function pointers forming Parrot's unified type interface. This interface is a union of numeric, string, array, hash and object like behaviors. For example the  \textit{get\_integer} function returns an integer value for the data type. For a simple \textit{int} it is the contained value. For an array it may be the number of elements. The \textit{find\_method} function makes user defined methods of objects possible.

@@

Control flow in Parrot is modeled using continuation passing. A continuation is a data structure containing the state of a program at a given point in its execution. It contains all information needed to continue a program in a certain state, e.g. a call stack, the instruction pointer and contents of local variables. ``Calling'' a continuation means restoring the state encapsulated in the continuation.

@@

A runloop is in principle the inner most loop executing bytecode in Parrot. It is also a data structure containing some information needed to support exception handling. When some operation in Parrot bytecode is calling a C function and this C function in turn is again executing Parrot bytecode, a nested runloop is started. Examples for such situations include calling a library function with a callback as parameter and exceptions thrown inside Parrot's C code calling a previously defined exception handler in user code.

@@

\textit{Parrot\_cx\_schedule\_sleep} is the actual implementation of the \textit{sleep} op. Like in \textit{Parrot\_cx\_preempt\_task}, \textit{Parrot\_cx\_stop\_task} is used to get an updated task at the current execution position, but instead of pushing this on the run queue, it is used as a callback for a newly set alarm. Again the NULL opcode is used to stop processing of the current task.
